\documentclass[../../../main.tex]{subfiles}
\begin{document}

\bibliographystyle{plainnat}
\begin{thebibliography}{99}

\bibitem[Baumslag and Chandler(1968)]{BaumslagAndChandler1968}
  Baumslag, B. and B. Chandler (1968).
  \emph{Group Theory} (Schaum's Outline Series).
  New York: McGRaw-Hill Book Company.
  
\bibitem[Benjamin et al.(2015)]{BenjaminEtAl2015}
  Benjamin, Arthur, Gary Chartrand, and Ping Zhang (2015).
  \emph{The Fascinating World of Graph Theory}.
  Princeton, NJ: Princeton University Press.

\bibitem[Blackburn and Bos(2005]{BlackburnAndBos2005}
  Blackburn, Patrick and Johan Bos (2005).
  \emph{Representation and Inference for Natural Language: A First Course in Computational Semantics}.
  Stanford, CA: CSLI Publications.
  
\bibitem[Boolos, Burgess, and Jeffrey(2002)]{BoolosBurgessAndJeffrey2002}
  Boolos, George, John Burgess, and Richard Jeffrey (2002).
  \emph{Computability and Logic}, 4th edition.
  Cambridge: Cambridge University Press.

\bibitem[Burger and Starbird(2010)]{BurgerAndStarbird2010}
  Burger, Edward and Michael Starbird (2010).
  \emph{The Heart of Mathematics: An Invitation to Effective Thinking}, 3rd ed.
  Hoboken, NJ: John Wiley \& Sons, Inc.

\bibitem[Chartrand(1977)]{Chartrand1977}
  Chartrand, Gary (1977).
  \emph{Introductory Graph Theory}.
  New York: Dover Publications, Inc. Reprint of the Prindle, Weber, \& Schmidt edition.
  
\bibitem[Cummings(2020)]{Cummings2020}
  Cummings, Jay (2020).
  \emph{Proofs: A Long-Form Mathematics Textbook}.
  Self-published.
  
\bibitem[Cummings(2018)]{Cummings2018}
  Cummings, Jay (2018).
  \emph{Real Analysis: A Long-Form Mathematics Textbook}, 2nd ed.
  Self-published.

\bibitem[Davey and Priestley(1990)]{DaveyAndPriestly1990}
  Davey, B. A. and H. A. Priestley (1990).
  \emph{Introduction to Lattices and Order}.
  Cambridge: Cambridge University Press.

\bibitem[Devlin(2012)]{Devlin2012}
  Devlin, Keith (2012).
  \emph{Introduction to Mathematical Thinking}.
  Petaluma, CA: Self-published by Keith Devlin.

\bibitem[Dos Reis and Dos Reis(2017)]{DosReisAndDosReis2017}
  Dos Reis, Laura L. and Anthony J. Dos Reis (2017).
  \emph{Abstract Algebra: A Student-Friendly Approach}.
  Self-published.
  
\bibitem[Dowty et al.(1981)]{DowtyEtAl1981}
  Dowty, David R., Robert E. Wall, and Stanley Peters (1981).
  \emph{Introduction to Montague Semantics}.
  Boston: Kluwer Academic Publishers.
  
\bibitem[Enderton(1977)]{Enderton1977}
  Enderton, Herbert.
  \emph{Elements of Set Theory}.
  Boston: Academic Press.
  
\bibitem[Flegg(1974)]{Flegg1974}
  Flegg, H. Graham (1974).
  \emph{From Geometry to Topology}.
  New York: Dover Publications, Inc. Reprint of the Crane, Russak, and Co. edition.
  
\bibitem[Fournier(2009)]{Fournier2009}
  Fournier, Jean-Claude (2009).
  \emph{Graph Theory and Applications}.
  Hoboken, NJ: Wiley.
  
\bibitem[Gould(2012)]{Gould2012}
  Gould, Ronald (2012).
  \emph{Graph Theory}.
  New York: Dover Publications, Inc. Reprint of the 1988 Benjamin-Cummings edition.
  
\bibitem[Halmos(2017)]{Halmos2017}
  Halmos, Paul.
  \emph{Naive Set Theory}.
  New York: Dover Publications, Inc. Reprint of the 1960 Van Nostrand edition.
  
\bibitem[Hartsfield and Ringel(1990)]{HartsfieldAndRingel1990}
  Hartsfield, Nora and Gerhard Ringel (1990).
  \emph{Pearls of Graph Theory: A Comprehensive Introduction}.
  Boston: Academic Press, Inc.
  
\bibitem[Jongsma(2019)]{Jongsma2019}
  Jongsma, Calvin (2019).
  \emph{Introduction to Discrete Mathematics via Logic and Proof}.
  Cham, Switzerland: Springer.

\bibitem[Madden and Aubrey(2017)]{MaddenAndAubrey2017}
  Madden, Daniel J. and Jason A. Aubrey (2017).
  \emph{An Introduction to Proof through Real Analysis}.
  Hoboken, NJ: Wiley.

\bibitem[Pinter(2014)]{Pinter2014}
  Pinter, Charles (2014).
  \emph{A Book of Set Theory}.
  New York: Dover Publications, Inc. Reprint of the 1971 Addison-Wesley edition.
  
\bibitem[Pinter(2010)]{Pinter2010}
  Pinter, Charles (2010).
  \emph{A Book of Abstract Algebra}, 2nd edition.
  New York: Dover Publications, Inc. Reprint of the 1982 McGraw-Hill edition (also reprinted by Dover in 1990).

\bibitem[Saracino(2008)]{Saracino2008}
  Saracino, Dan (2008).
  \emph{Abstract Algebra}, 2nd edition.
  Long Grove, IL: Waveland Press, Inc.

\bibitem[Sawyer(1982)]{Sawyer1982}
  Sawyer, W. W. (1982).
  \emph{Prelude to Mathematics}.
  New York: Dover Publications, Inc. An enlarged version of the 1955 Penguin edition.

\bibitem[Sawyer(2007)]{Sawyer2007}
  Sawyer, W. W. (2007).
  \emph{Mathematician's Delight}.
  New York: Dover Publications, Inc. Reprint of the 1943 Penguin edition (also reprinted by Dover in 1991.

\bibitem[Sawyer(2018)]{Sawyer2018}
  Sawyer, W. W. (2018).
  \emph{A Concrete Approach to Abstract Algebra}.
  New York: Dover Publications, Inc. Reprint of the 1959 W. H. Freeman and Company edition (also reprinted by Dover in 1978).

\bibitem[Schmidt(1966)]{Schmidt1966}
  Schmidt, O. U. (1966).
  \emph{Abstract Theory of Groups}, trans. by Fred Holling and J. B. Roberts, ed. by J. B. Roberts.
  San Francisco, CA: W. H. Freeman and Company.

\bibitem[Smullyan and Fitting(2010)]{SmullyanAndFitting2010}
  Smullyan, Raymond and Melvin Fitting (2010).
  \emph{Set Theory and the Continuum Problem}.
  New York: Dover Publications, Inc. Reprint of the 1996 Oxford University Press edition.

\bibitem[Steinhart(2018)]{Steinhart2018}
  Steinhart, Eric (2018).
  \emph{More Precisely: The Math You Need to Do Philosophy}, 2nd ed.
  Peterborough, Ontario: Broadview Press.

\bibitem[Stewart(1995)]{Stewart1995}
  Stewart, Ian (1995).
  \emph{Concepts of Modern Mathematics}.
  New York: Dover Publications, Inc. Reprint of the 1981 Penguin edition.

\bibitem[Stewart and Tall (2015)]{StewartAndTall2015}
  Stewart, Ian and David Tall (2015).
  \emph{The Foundations of Mathematics}, 2nd. edition.
  Oxford: Oxford University Press.

\bibitem[Teller(1989)]{Teller1989}
  Teller, Paul (1989).
  \emph{A Modern Formal Logic Primer}, 2 volumes.
  Upper Saddle River, NJ: Prentice Hall.

\bibitem[Trudeau(1993]{Trudeau1993}
  Trudeau, Richard (1993).
  \emph{Introduction to Graph Theory}.
  New York: Dover Publications, Inc. Reprint of the 1976 Kent State University Press edition.

\bibitem[Velleman(2019)]{Velleman2019}
  Velleman, Daniel (2019).
  \emph{How to Prove It: A Structured Approach}, 3rd edition.
  Cambridge: Cambridge University Press.

\bibitem[Warner(1990)]{Warner1990}
  Warner, Seth (1990).
  \emph{Modern Algebra}, two volumes bound as one.
  New York: Dover Publications, Inc. Reprint of the 1965 Prentice-Hall 2-volume edition.

\bibitem[Warner(2019)]{Warner2019}
  Warner, Steve (2019).
  \emph{Topology for Beginners}.
  Self-published.

\bibitem[Wilder(2012)]{Wilder2012}
  Wilder, Raymond (2012).
  \emph{Introduction to the Foundations of Mathematics}, 2nd edition.
  New York: Dover Publications, Inc. Reprint of the 1965 John Wiley \& Sons edition (also reprinted by Dover in 1983).

\bibitem[Wolf(1998)]{Wolf1998}
  Wolf, Robert (1998).
  \emph{Proof, Logic, and Conjecture: The Mathematician's Toolbox}.
  New York: W.H. Freeman and Company.

\bibitem[Zach(2019)]{Zach2019}
  Zach, Richard (2019).
  \emph{Sets, Logic, Computation: An Open Introduction to Metalogic}, Fall 2019 edition.
  Self-published.

\end{thebibliography}


\end{document}
