\documentclass[../../../main.tex]{subfiles}
\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Further Reading}

To pursue functions further, the following list may offer some helpful starting points.

\begin{itemize}

  \item \citet[ch.~5]{StewartAndTall2015} provides a good but introductory-level discussion to functions.
  
  \item \citet[ch.~2]{Steinhart2018} also provides an excellent introduction for the beginner to the topic of functions (and relations).
  
  \item \citet[ch.~5]{Stewart1995} offers a discussion of functions for non-experts.
  
  \item \citet[ch.~15]{Flegg1974} provides a conceptual discussion of functions.
  
  \item \citet[ch.~8]{Cummings2020} presents a slow, thorough discussion of functions and how to construct proofs about them.
  
  \item \citet[ch.~3]{Zach2019} offers a nice, introductory (yet still technical) discussion of functions.

  \item \citet[ch.~5]{Velleman2019} provides an excellent introduction to the concept of functions and proofs about them.

  \item \citet[chs.~11--12]{MaddenAndAubrey2017} give thorough coverage of functions and proofs about them.

  \item \citet[ch.~4]{Warner2019} provides thorough coverage of functions and proofs about them.

  \item \citet[ch.~2]{Pinter2014} offers a simple discussion of many basic theorems and proofs about functions.

  \item \citet[ch.~6]{Jongsma2019} provides a rigorous introduction to some of the basic theorems and proofs about functions (and equivalence classes).
  
\end{itemize}

\end{document}