\documentclass[../../../main.tex]{subfiles}
\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Further Reading}

To pursue sets and set theory further, the following list may offer some helpful starting points.

\begin{itemize}

  \item \citet[ch.~3]{StewartAndTall2015} provides a good but introductory-level discussion to the basic concept of set theory.
  
  \item \citet[ch.~1]{Steinhart2018} also provides an excellent introduction for the beginner to the topic of sets and set theory.
  
  \item \citet[ch.~4]{Stewart1995} offers a discussion of set theory for non-experts.
  
  \item \citet[ch.~14]{Flegg1974} provides a conceptual discussion of sets and set theory.

  \item \citet{Halmos2017} offers a non-technical discussion of the classic topics in set theory.
  
  \item \citet[ch.~3]{Cummings2020} presents a slow, thorough discussion of sets and how to construct proofs about them.

  \item \citet[chs.~1--2]{Velleman2019} discusses logic in these chapters, but in the midst of all that, he provides an excellent introduction to the concept of sets and proofs about them.
  
  \item \citet[ch.~1]{Zach2019} provides an introductory-level discussion of sets, with proofs and some helpful discussion.
  
  \item \citet[chs.~1--2]{Warner2019} covers sets in a thorough way, with worked out proofs for all details.
  
  \item \citet[ch.~9]{MaddenAndAubrey2017} provides a discussion of sets and how to build proofs about them.

  \item \citet[ch.~1]{Pinter2014} offers a simple discussion of the basic theorems and proofs for set theory.

  \item \citet[chs.~4--5]{Jongsma2019} provides a rigorous introduction to many of the higher-level topics of set theory.
  
  \item \citet{Enderton1977} is a classic textbook for set theory. If you find set theory deeply fascinating, turn to this after you are comfortable reading proofs.
  
  \item \citet{SmullyanAndFitting2010} is another textbook that thoroughly covers most of the classic topics in modern set theory. Reach for this one after you are comfortable reading proofs, and if you are interested in the many deep and interesting questions that can be found in set theory.

\end{itemize}

\end{document}